### 2014-10-08 paper.tex

  Herein we share a data set of typical gait data for 11 subjects walking at a
  three speeds on an instrumented treadmill while being longitudinally
  perturbed in the stance phase with psuedo-random fluctuations in the
  treadmill belt speed. We provide raw marker and ground reaction loads in
  addition to processed data that includes gait landmarks, 2D joint angles,
  angular rates, and joint torques. The protocol is described in detail along
  with the the additional meta data about each of the data files. This data can
  be useful for validating or genreating mathematical models that capable of
  simulating non-periodic and pertubed gaits.

@@

\section{Introduction}
%
Emergig powered prothetics are in need of bio-inspired control systems.
Ideally, a powered prosthetic would behave in such a way that the user would
feel like their limb was never disabled. There are a variety of approaches to
developing bio-inspired control systems some of which aim to mimic the
reactions and motion of an abled bodied person. A modern gait lab is able to
collect a variety of kinmetic, kinetic, and physiological data from humans
during gait. This data can be used to drive the design of the controller. With
a rich enough data set, one may be able to identify control mechanism used
during a human's natural gait and recovery from perturbations. We have
collected one such data set that we believe is richer than previous gait data
sets and may be rich enough for control identification. The data can also be
used for verifciation purposes for controllers that have been designed in other
manners.

@@

The output from the D-Flow mocap module is stored in a tab delimted file named
\verb+mocap-xxx.txt+ where \verb+xxx+ represents the trial id number. The file
contains several columns:

\begin{description}
  \item[TimeStamp] The monotonically increasing computer clock time when D-Flow
    recieves a frame from Cortex. These are approximately at 100 hz and given
    in seconds.
  \item[FrameNumber] Monotonically increasing positive integers that correspond
    to each frame received from Cortex.
  \item[Marker Coordinates] Any column that ends in \verb+.PosX+, \verb+.PosY+,
    or \verb+.PosZ+ are marker coordinates expressed in Cortex's Cartesian
    reference frame. These values are in meters.
  \item[Ground Reaction Loads] There are three ground reaction forces and three
    ground reaction moments recorded by each of the two force plates in Newtons
    and Newton-Meters, respectively. The prefix for these columns is either
    \verb+FP1+ or \verb+FP2+ and represents either force plate 1 (left) or 2
    (right). The suffixes are either \verb+.For[XYZ]+, \verb+.Mom[XYZ]+ for the
    forces and moments, respectively. The force plate voltages are sampled at a
    much higher frequency than the cameras, but delivered at the Cortex camera
    sample rate, 100 Hz through the D-Flow mocap module. A force/moment
    calibration matrix stored in Cortex converts the voltages to forces and
    moments before sending it to D-Flow [1]. Cortex also computes the center of
    pressure from the forces, moments, and force plate dimensions. These have
    the same prefixes for the plate number, have the suffix \verb+.Cop[XYZ]+,
    and are in meters.
  \item[Analog Channels] Several analog signals are recorded under column
    headers \verb+Channel[1-99].Anlg+. These correspond to analog signals
    sampled by Cortex and correspond to the ?92? National Instruments XXXX DAQ
    box. The first twelve are the voltages from the force plate load cells. We
    also record the acceleration of 4 points on the treadmill base in chanels
    X-X.
\end{description}

@@

The \verb+study+ section contains identififying information for the overall
study, an identification number, name, and description. This is the same for
all meta data files in the study.

@@

>>>>

  Herein we share a data set of typical gait data for 11 subjects walking at a
  three speeds on an instrumented treadmill while being longitudinally
  perturbed in the stance phase with pseudorandom fluctuations in the
  treadmill belt speed. We provide raw marker and ground reaction loads in
  addition to processed data that includes gait landmarks, 2D joint angles,
  angular rates, and joint torques. The protocol is described in detail along
  with the the additional meta data about each of the data files. This data can
  be useful for validating or genreating mathematical models that capable of
  simulating non-periodic and pertubed gaits.

@@

\section{Introduction}
%
Emerging powered prothetics are in need of bio-inspired control systems.
Ideally, a powered prosthetic would behave in such a way that the user would
feel like their limb was never disabled. There are a variety of approaches to
developing bio-inspired control systems some of which aim to mimic the
reactions and motion of an able-bodied person. A modern gait lab is able to
collect a variety of kinmetic, kinetic, and physiological data from humans
during gait. This data can be used to drive the design of the controller. With
a rich enough data set, one may be able to identify control mechanism used
during a human's natural gait and recovery from perturbations. We have
collected one such data set that we believe is richer than previous gait data
sets and may be rich enough for control identification. The data can also be
used for verifciation purposes for controllers that have been designed in other
manners.

@@

The output from the D-Flow mocap module is stored in a tab delimited file named
\verb+mocap-xxx.txt+ where \verb+xxx+ represents the trial id number. The file
contains several columns:

\begin{description}
  \item[TimeStamp] The monotonically increasing computer clock time when D-Flow
    recieves a frame from Cortex. These are approximately at 100 hz and given
    in seconds.
  \item[FrameNumber] Monotonically increasing positive integers that correspond
    to each frame received from Cortex.
  \item[Marker Coordinates] Any column that ends in \verb+.PosX+, \verb+.PosY+,
    or \verb+.PosZ+ are marker coordinates expressed in Cortex's Cartesian
    reference frame. These values are in meters.
  \item[Ground Reaction Loads] There are three ground reaction forces and three
    ground reaction moments recorded by each of the two force plates in Newtons
    and Newton-Meters, respectively. The prefix for these columns is either
    \verb+FP1+ or \verb+FP2+ and represents either force plate 1 (left) or 2
    (right). The suffixes are either \verb+.For[XYZ]+, \verb+.Mom[XYZ]+ for the
    forces and moments, respectively. The force plate voltages are sampled at a
    much higher frequency than the cameras, but delivered at the Cortex camera
    sample rate, 100 Hz through the D-Flow mocap module. A force/moment
    calibration matrix stored in Cortex converts the voltages to forces and
    moments before sending it to D-Flow [1]. Cortex also computes the center of
    pressure from the forces, moments, and force plate dimensions. These have
    the same prefixes for the plate number, have the suffix \verb+.Cop[XYZ]+,
    and are in meters.
  \item[Analog Channels] Several analog signals are recorded under column
    headers \verb+Channel[1-99].Anlg+. These correspond to analog signals
    sampled by Cortex and correspond to the 96 analog-channel National 
    Instruments 6255 DAQ box. The first twelve are the voltages from the 
    force plate load cells. We also record the acceleration of 4 points on the
    treadmill base in channels 61-72.
\end{description}

@@

The \verb+study+ section contains identifying information for the overall
study, an identification number, name, and description. This is the same for
all meta data files in the study.

@@

\section{Perturbation Signals}

  The data set uses a testing protocol containing a segment of normal 
  walking, followed by a series of pseudorandom longitudinal 
  perturbations, and ending with a second segment of normal walking.   
  Three pseudorandom speed control signals, with mean velocities of 
  0.8, 1.2, and 1.6 m/s, were generated through MATLAB and Simulink. 
  Initially, a random acceleration signal was created using discrete-time 
  Gaussian white noise and saturated at the maximum belt acceleration 
  of 15 m/s2, integrated to obtain belt speed, and high-pass filtered 
  (second-order Butterworth) to eliminate velocity drift.   Mean velocity was 
  then added to the signal and limited between 0 and 3.6 m/s.  The cutoff
  frequencies of the high-pass filter, as well as the variance in the 
  acceleration signal, were adjusted until the appropriate standard deviations
  for each mean velocity were obtained.

  These standard deviations were determined through a trial and error method.
  Initially, perturbations of $\pm$ 10% for each mean velocity were generated. 
  However, this large deviation in the slowest mean velocity (0.8 m/s) would
  unintentionally reduce the treadmill speed to values close to zero.  This 
  would produce an undesired stopping and starting sensation.  Alternatively, 
  smaller standard perturbations of $\pm$ 5% did not produce noticeable 
  perturbations in the fastest trial (1.6 m/s).  Therefore, standard deviations
  of 7.16%, 10.23%, and 13.1% were selected for the 0.8, 1.2, and 
  1.6 m/s speeds, respectively.   This ensured that the test subject was
  sufficiently perturbed for each speed, while remaining within the limits of 
  our testing protocol.

  To ensure that the treadmill could follow the higher frequencies of the 
  random signals, the maximum treadmill acceleration of 15 m/s2 was used in the
  D-Flow Software.  As this value remains constant throughout the experiment,
  the initial starting and stopping of the treadmill would be too rapid for test
  subjects.  To eliminate this, a ramped acceleration and deceleration was 
  generated and placed at the beginning and end of the entire signal,
  respectively.   Other additions include the one minute segments of 
  unperturbed walking at the beginning and end of the trial and 30 seconds of 
  no movement for force plate zeroing and subject T-pose calibration.  The 
  sampling rate of the simulation was set to 100 Hz, corresponding to the 
  D-Flow software's capabilities.

  The resulting output from the MATLAB script and Simulink produces a comma-
  delimited text file of six columns: timestamp, slow, normal, and fast walking
  perturbation signals, and slow and fast running signals.  These running 
  signals were generated in the event the testing protocol expanded from a
  walking study to a running study, though they were not used in the data set.
